Bayesian Shape Reconstruction and Optimal Guidance for Autonomous Landing on Asteroids

### 2019-02-15 manuscript.tex
Add link to videos
Updating manuscript

\section*{Nomenclature}

\noindent(Nomenclature entries should have the units identified)
\noindent{Add elements}
% {\renewcommand\arraystretch{1.0}
% \noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
% $A$  & amplitude of oscillation \\
% $a$ &    cylinder diameter \\
% $C_p$& pressure coefficient \\
% $Cx$ & force coefficient in the \textit{x} direction \\
% $Cy$ & force coefficient in the \textit{y} direction \\
% c   & chord \\
% d$t$ & time step \\
% $Fx$ & $X$ component of the resultant pressure force acting on the vehicle \\
% $Fy$ & $Y$ component of the resultant pressure force acting on the vehicle \\
% $f, g$   & generic functions \\
% $h$  & height \\
% $i$  & time index during navigation \\
% $j$  & waypoint index \\
% $K$  & trailing-edge (TE) nondimensional angular deflection rate\\
% $\Theta$ & boundary-layer momentum thickness\\
% $\rho$ & density\\
% \multicolumn{2}{@{}l}{Subscripts}\\
% cg & center of gravity\\
% $G$ & generator body\\
% iso	& waypoint index
% \end{longtable*}}

@@

\todo{Define every variable, $m_1,m_2,\rho_1,\ldots$. }

@@

We represent the gravitational potential of the asteroid using a polyhedron gravitation model.
This model is composed of a polyhedron, which is a three-dimensional solid body, that is defined by a series of vectors in the body-fixed frame.
The vectors define vertices in the body-fixed frame as well as planar faces which compose the surface of the asteroid.
We assume that each face is a triangle composed of three vertices and three edges.
As a result, only two faces meet at each edge while three faces meet at each vertex.
Only the body-fixed vectors, and their associated topology, is required to define the exterior gravitational model.
References~\cite{werner1994} and~\cite{werner1996} give a detailed derivation of the polyhedron model.
Here, we summarize the key equations required for implementation.

We define the attraction, gravity gradient matrix, and Laplacian as
\begin{align}
    \nabla U ( \vc{r} ) &= -G \sigma \sum_{e \in \text{edges}} \vc{E}_e \cdot \vc{r}_e \cdot L_e + G \sigma \sum_{f \in \text{faces}} \vc{F}_f \cdot \vc{r}_f \cdot \omega_f \in \R^{3 \times 1} , \label{eq:attraction}\\
    \nabla \nabla U ( \vc{r} ) &= G \sigma \sum_{e \in \text{edges}} \vc{E}_e  \cdot L_e - G \sigma \sum_{f \in \text{faces}} \vc{F}_f \cdot \omega_f \in \R^{3 \times 3}, \label{eq:gradient_matrix}\\
    \nabla^2 U &= -G \sigma \sum_{f \in \text{faces}}  \omega_f \in \R^1 .\label{eq:laplacian}
\end{align}
One interesting thing to note is that both~\cref{eq:face_dyad,eq:edge_dyad} can be precomputed without knowledge of the position of the satellite.
They are both solely functions of the vertices and edges of the polyhedral shape model and are computed once and stored.
Once a position vector \( \vc{r} \) is defined, the scalars given in~\cref{eq:edge_factor,eq:face_factor} can be computed for each face and edge.
Finally,~\cref{eq:potential} is used to compute the gravitational potential on the spacecraft.
The Laplacian, defined in~\cref{eq:laplacian}, gives a simple method to determine if the spacecraft has collided with the body~\cite{werner1996}.

@@

First, an attitude error function \(A : \SO \times \SO  \to \R \), an attitude error vector \( e_R : \SO \times \SO \to \R \), and an angular velocity error \( e_\Omega : \SO \times \R^3 \times \SO \times \R^3 \) are defined as
\begin{subequations}\label{eq:attitude_error_function}
\begin{align}
    A(R, R_d) &= \frac{1}{2}  \tr{G\parenth{I - R_d^T R}}, \label{eq:A} \\
    e_R &= \frac{1}{2} G \parenth{R_d^T R - R^T R_d^\vee}, \label{eq:eRA}\\
    e_\Omega &= \Omega - R^T R_d \Omega_d \label{eq:eWA}.
\end{align}
\end{subequations}
Then the following properties hold:
\begin{enumerate}
    \item \label{item:prop_A_psd} \( A \) is  positive definite about \( R = R_d \) on \( \SO \).
    \item \label{item:prop_eRA} The variation of \( A \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}\label{eq:dirDiff_A}
		\dirDiff{A}{R} &= \eta \cdot e_{R} ,
	\end{align}
	where the notation \( \dirDiff{A}{R} \) represents the directional derivative of $A$ with respect to $R$ along the direction $\delta R$.
\item \label{item:prop_A_critical_points} The critical points of \( A \), where \( e_R = 0 \) are \( \braces{R_d} \cup \braces{R_d \exp(\pi \hat s) } \) for \( s \in \braces{e_1, e_2, e_3}\)
    \item \label{item:prop_A_quadratic} \( \Psi \) is a locally quadratic function, which means there exist constants \( 0 < n_1 \leq n_2 \) such that
    \begin{align}\label{eq:A_bound}
        n_1 \norm{e_R}^2 \leq \Psi(R) \leq n_2 \norm{e_R}^2 ,
    \end{align}
    for $0<\psi < h_1 $ where the constants \( n_1 = \frac{h_1}{h_2 + h_3} \) and \( n_2 = \frac{h_1 h_4}{h_5 \parenth{h1 - \psi} }\) for
	\begin{align*}
		h_1 &= \min\braces{g_1 + g_2, g_2 + g_3 , g_3 + g_1} ,\\
		h_2 &= \max\braces{\parenth{g_1 -g_2}^2,\parenth{g_2 -g_3}^2 , \parenth{g_3 -g_1}^2} ,\\
		h_3 &= \max\braces{\parenth{g_1 + g_2}^2, \parenth{g_2 + g_3}^2 , \parenth{g_3 + g_1}^2}, \\		
        h_4 &= \max\braces{g_1 + g_, g_2 +g_3, g_3 + g_1}, \\
        h_5 &= \max\braces{\parenth{g_1 + g_2}^2, \parenth{g_2 + g_3}^2, \parenth{g_3 + g_1}^2}.
	\end{align*}
\end{enumerate}
The proof of these properties is available~\cref{proof:attitude_configuration_error}.

With the appropriate attitude configuration error we now present the error dynamics in~\cref{prop:attractive_error_dynamics}, which are used in the subsequent development of the nonlinear control system.
The attitude error dynamics for \( A(R, R_d), e_\Omega \) satisfy
\begin{gather}
    \diff{}{t} \parenth{A(R)} = e_{R_A} \cdot e_\Omega , \label{eq:A_dot} \\
    \diff{}{t} \parenth{e_{R_A}} = E(R, R_d) e_\Omega , \label{eq:eRA_dot} \\
    \diff{}{t} \parenth{e_\Omega} = J^{-1} \parenth{-\hat{\Omega} J \Omega + u_m + M_{ext}} , \label{eq:eW_dot}
\end{gather}
where the matrix \(E(R,R_d)\in \R^{3\times3} \) is given by
\begin{align}
    E(R,R_d) = &\frac{1}{2} \parenth{\tr{R^T R_d G}I - R^T R_d G} , \label{eq:E} \\
\end{align}
The proof is show in~\Cref{proof:attractive_error_dynamics}
With the appropriate attitude configuration error we now present an attitude controller for attitude stabilization and tracking of the inertial attitude dynamics given in~\cref{eq:inertial_attitude_dynamics,eq:inertial_angvel_dynamics}.

Given a desired attitude command \( \parenth{R_d, \Omega_d = 0} \) and positive constants \( k_R, k_\Omega \in \R \) we define a control input \( u \in \R^3 \) as follows
\begin{gather}
    u = -k_R e_R - k_\Omega e_\Omega + \Omega \times J \Omega - M_{ext}. \label{eqn:nodist_control}
\end{gather}
Then the zero equilibrium of the attitude error is asymptotically stable.
The proof is show in~\cref{proof:attitude_controller}.

The required translational control are defined in terms of the position and velocity tracking errors.
The tracking error vectors are easier to define as they evolve on a Euclidean space rather than a nonlinear manifold and are given by
\begin{subequations}\label{eq:translation_error_variables}
\begin{align}
    e_x &= x - x_d ,\\
    e_v &= v - \dot{x}_d.
\end{align}
\end{subequations}
The error dynamics are given by
\begin{subequations}\label{eq:translation_error_dynamics}
\begin{align}
    \dot{e}_x &= v - \dot{x}_d, \\
    \dot{e}_v &= \dot{v} - \ddot{x}_d .
\end{align}
\end{subequations}
A control input, \( u_f \), is derived to ensure asymptotic trajectory tracking for the translational dynamics.
    Given a desired trajectory \( \vc{x}_d \) and positive constants \( k_x, k_v, c \in \R \) we define a control input \( \vc{u}_f \in \R \) as follows:
    \begin{align}\label{eq:translational_control}
        \vc{u}_f = -k_x e_x - k_v e_v - F_{ext} + m \ddot{x}_d .
    \end{align}
    If \( c \) is chosen such that
    \begin{align}\label{eq:translational_control_gain_bound}
        0 < c < \min \braces{\sqrt{k_x m}, \frac{4 k_x k_v m}{ k_v \parenth{4 k_x m - k_v} }   }
    \end{align}
    then the zero equilibrium of the error vectors~\cref{eq:translation_error_variables} is stable in the sense of Lyapunov.
    Furthermore, \( e_x, e_v \to 0 \) as \( t \to \infty\).
The proof is shown in~\cref{proof:translational_control}.

@@

The final phase of the simulation is to utilize the estimated shape model and vertically descend to the desired landing site. 
This is accomplished using the closed loop control of the vehicle and a trajectory which transitions from the home position to the surface over \SI{3600}{\second}.
The landing trajectory, using the estimated shape, is visualized in the asteroid frame in~\cref{fig:castalia_landing}.
During the vertical descent the vehicle is able to accurately track the desired trajectory using the estimated shape to compute the gravitational potential.
\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/dynamic_exploration/castalia/land/asteroid_trajectory.jpg}
    \caption{Vertical descent onto 4769 Castalia~\label{fig:castalia_landing}}
\end{figure}

\todo{as the nonlinear controller is added, we may want to add more plots for tracking errors; we can add a link to videos}

>>>>

% \section*{Nomenclature}

% \noindent(Nomenclature entries should have the units identified)
% \noindent{Add elements}
% {\renewcommand\arraystretch{1.0}
% \noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
% $A$  & amplitude of oscillation \\
% $a$ &    cylinder diameter \\
% $C_p$& pressure coefficient \\
% $Cx$ & force coefficient in the \textit{x} direction \\
% $Cy$ & force coefficient in the \textit{y} direction \\
% c   & chord \\
% d$t$ & time step \\
% $Fx$ & $X$ component of the resultant pressure force acting on the vehicle \\
% $Fy$ & $Y$ component of the resultant pressure force acting on the vehicle \\
% $f, g$   & generic functions \\
% $h$  & height \\
% $i$  & time index during navigation \\
% $j$  & waypoint index \\
% $K$  & trailing-edge (TE) nondimensional angular deflection rate\\
% $\Theta$ & boundary-layer momentum thickness\\
% $\rho$ & density\\
% \multicolumn{2}{@{}l}{Subscripts}\\
% cg & center of gravity\\
% $G$ & generator body\\
% iso	& waypoint index
% \end{longtable*}}

@@

@@

We represent the gravitational potential of the asteroid using a polyhedron gravitation model.
This model is composed of a polyhedron, which is a three-dimensional solid body, that is defined by a series of vectors in the body-fixed frame.
The vectors define vertices in the body-fixed frame as well as planar faces which compose the surface of the asteroid.
We assume that each face is a triangle composed of three vertices and three edges.
As a result, only two faces meet at each edge while three faces meet at each vertex.
Only the body-fixed vectors, and their associated topology, is required to define the exterior gravitational model.
References~\cite{werner1994} and~\cite{werner1996} give a detailed derivation of the polyhedron model.
Here, we summarize the key equations required for implementation.
We define the attraction, gravity gradient matrix, and Laplacian as
\begin{align}
    \nabla U ( \vc{r} ) &= -G \sigma \sum_{e \in \text{edges}} \vc{E}_e \cdot \vc{r}_e \cdot L_e + G \sigma \sum_{f \in \text{faces}} \vc{F}_f \cdot \vc{r}_f \cdot \omega_f \in \R^{3 \times 1} , \label{eq:attraction}\\
    \nabla \nabla U ( \vc{r} ) &= G \sigma \sum_{e \in \text{edges}} \vc{E}_e  \cdot L_e - G \sigma \sum_{f \in \text{faces}} \vc{F}_f \cdot \omega_f \in \R^{3 \times 3}, \label{eq:gradient_matrix}\\
    \nabla^2 U &= -G \sigma \sum_{f \in \text{faces}}  \omega_f \in \R^1 .\label{eq:laplacian}
\end{align}
The Laplacian, defined in~\cref{eq:laplacian}, gives a simple method to determine if the spacecraft has collided with the body~\cite{werner1996}.

@@

First, an attitude error function \(A : \SO \times \SO  \to \R \), an attitude error vector \( e_R : \SO \times \SO \to \R \), and an angular velocity error \( e_\Omega : \SO \times \R^3 \times \SO \times \R^3 \) are defined as
\begin{subequations}\label{eq:attitude_error_function}
\begin{align}
    A(R, R_d) &= \frac{1}{2}  \tr{G\parenth{I - R_d^T R}}, \label{eq:A} \\
    e_R &= \frac{1}{2} G \parenth{R_d^T R - R^T R_d^\vee}, \label{eq:eRA}\\
    e_\Omega &= \Omega - R^T R_d \Omega_d \label{eq:eWA}.
\end{align}
\end{subequations}
Then the following properties hold:
\begin{enumerate}
    \item \label{item:prop_A_psd} \( A \) is  positive definite about \( R = R_d \) on \( \SO \).
    \item \label{item:prop_eRA} The variation of \( A \) with respect to a variation of \( \delta R = R \hat{\eta} \) for \( \eta \in \R^3 \) is given by
	\begin{align}\label{eq:dirDiff_A}
		\dirDiff{A}{R} &= \eta \cdot e_{R} ,
	\end{align}
	where the notation \( \dirDiff{A}{R} \) represents the directional derivative of $A$ with respect to $R$ along the direction $\delta R$.
\item \label{item:prop_A_critical_points} The critical points of \( A \), where \( e_R = 0 \) are \( \braces{R_d} \cup \braces{R_d \exp(\pi \hat s) } \) for \( s \in \braces{e_1, e_2, e_3}\)
    \item \label{item:prop_A_quadratic} \( \Psi \) is a locally quadratic function, which means there exist constants \( 0 < n_1 \leq n_2 \) such that
    \begin{align}\label{eq:A_bound}
        n_1 \norm{e_R}^2 \leq \Psi(R) \leq n_2 \norm{e_R}^2 ,
    \end{align}
    for $0<\psi < h_1 $ where the constants \( n_1 = \frac{h_1}{h_2 + h_3} \) and \( n_2 = \frac{h_1 h_4}{h_5 \parenth{h1 - \psi} }\) for
	\begin{align*}
		h_1 &= \min\braces{g_1 + g_2, g_2 + g_3 , g_3 + g_1} ,\\
		h_2 &= \max\braces{\parenth{g_1 -g_2}^2,\parenth{g_2 -g_3}^2 , \parenth{g_3 -g_1}^2} ,\\
		h_3 &= \max\braces{\parenth{g_1 + g_2}^2, \parenth{g_2 + g_3}^2 , \parenth{g_3 + g_1}^2}, \\		
        h_4 &= \max\braces{g_1 + g_, g_2 +g_3, g_3 + g_1}, \\
        h_5 &= \max\braces{\parenth{g_1 + g_2}^2, \parenth{g_2 + g_3}^2, \parenth{g_3 + g_1}^2}.
	\end{align*}
\end{enumerate}
The proof of these properties is available~\cite{kulumani2017a}.

With the appropriate attitude configuration error we now present the error dynamics, which are used in the subsequent development of the nonlinear control system.
The attitude error dynamics for \( A(R, R_d), e_\Omega \) satisfy
\begin{gather}
    \diff{}{t} \parenth{A(R)} = e_{R_A} \cdot e_\Omega , \label{eq:A_dot} \\
    \diff{}{t} \parenth{e_{R_A}} = E(R, R_d) e_\Omega , \label{eq:eRA_dot} \\
    \diff{}{t} \parenth{e_\Omega} = J^{-1} \parenth{-\hat{\Omega} J \Omega + u_m + M_{ext}} , \label{eq:eW_dot}
\end{gather}
where the matrix \(E(R,R_d)\in \R^{3\times3} \) is given by
\begin{align}
    E(R,R_d) = &\frac{1}{2} \parenth{\tr{R^T R_d G}I - R^T R_d G} , \label{eq:E}
\end{align}
The proof is available in~\cite{kulumani2017a}.

Using these properties we can now define an attitude control input to allow the vehicle to track attitude errors.
Given a desired attitude command \( \parenth{R_d, \Omega_d = 0} \) and positive constants \( k_R, k_\Omega \in \R \) we define a control input \( u \in \R^3 \) as follows
\begin{gather}
    u = -k_R e_R - k_\Omega e_\Omega + \Omega \times J \Omega - M_{ext}. \label{eqn:nodist_control}
\end{gather}
Then the zero equilibrium of the attitude error is asymptotically stable.
The proof is show in~\cite{kulumani2017a}.

The required translational control are defined in terms of the position and velocity tracking errors.
The tracking error vectors are easier to define as they evolve on a Euclidean space rather than a nonlinear manifold and are given by
\begin{subequations}\label{eq:translation_error_variables}
\begin{align}
    e_x &= x - x_d ,\\
    e_v &= v - \dot{x}_d.
\end{align}
\end{subequations}
The error dynamics are given by
\begin{subequations}\label{eq:translation_error_dynamics}
\begin{align}
    \dot{e}_x &= v - \dot{x}_d, \\
    \dot{e}_v &= \dot{v} - \ddot{x}_d .
\end{align}
\end{subequations}
A control input, \( u_f \), is derived to ensure asymptotic trajectory tracking for the translational dynamics.
    Given a desired trajectory \( \vc{x}_d \) and positive constants \( k_x, k_v, c \in \R \) we define a control input \( \vc{u}_f \in \R \) as follows:
    \begin{align}\label{eq:translational_control}
        \vc{u}_f = -k_x e_x - k_v e_v - F_{ext} + m \ddot{x}_d .
    \end{align}
    If \( c \) is chosen such that
    \begin{align}\label{eq:translational_control_gain_bound}
        0 < c < \min \braces{\sqrt{k_x m}, \frac{4 k_x k_v m}{ k_v \parenth{4 k_x m - k_v} }   }
    \end{align}
    then the zero equilibrium of the error vectors~\cref{eq:translation_error_variables} is stable in the sense of Lyapunov.
    Furthermore, \( e_x, e_v \to 0 \) as \( t \to \infty\).
    The proof is shown in~\cite{kulumani2017a}.

@@

The final phase of the simulation is to utilize the estimated shape model and vertically descend to the desired landing site. 
This is accomplished using the closed loop control of the vehicle and a trajectory which transitions from the home position to the surface over \SI{3600}{\second}.
The landing trajectory, using the estimated shape, is visualized in the asteroid frame in~\cref{fig:castalia_landing}.
During the vertical descent the vehicle is able to accurately track the desired trajectory using the estimated shape to compute the gravitational potential.
\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/dynamic_exploration/castalia/land/asteroid_trajectory.jpg}
    \caption{Vertical descent onto 4769 Castalia~\label{fig:castalia_landing}}
\end{figure}
A number of videos are available at \url{http://bit.ly/shape_reconstruction} which demonstrate these results.
% \todo{as the nonlinear controller is added, we may want to add more plots for tracking errors; we can add a link to videos}
